Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2018-04-26 analysis/paper/paper.Rmd
Add a bit more to the discussion and cross references to figures and table 1

The plot design was a split-plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split-plot. The main plot treatments were alternate wetting and drying (AWD) and continuously flooded (FLD) or farmers' practice as the control treatment.

The main plot size was 12m x 12m (144 sq m), with a split-plot size of 5m x 5m (25 sq m). Replication size was 12m x 24m (288 sq m) with a buffer of 1m per split-plot for a whole experiment size of 1,152 sq m.

@@

The split-plot treatment rates of nitrogen were N0 (no nitrogen supply), N100 (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N120 (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation).

@@

Split-plot N rates differed from the 2015 rates with only two nitrogen treatments, N60 (60 kg per ha as urea split into two applications) and N180 (180 kg per ha in three splits) being applied.

@@

In both experiments we were unable to detect any significant effect of AWD on sheath blight that would cause increases in the disease, which could hinder adoption of the technology. In 2016 one measurement, tiller sheath blight severity, was significantly lower for AWD (FIGURE XXX) than the traditional FLD treatment, which may indicate a possible adverse effect of using AWD on leaf ShB severity.

The changes made for the 2016 experiment appear to have worked, increasing the incidence and decreased variability of sheath blight in the plots.

@@

```{r table-1, tab.cap="Nitrogen application rates for 2015 and 2016. Five different nitrogen treatments were applied depending on the experiment. In the 2015 experiment three nitrogen rate treatments were applied: no N (N0), 100 kg per ha N (N100) and 180 kg per ha N (N120). In the 2016 experiment, two nitrogen rate treatments were applied: 60 kg per ha N (N60) and 180 kg per ha N (N180). Treatments were applied in splits, basal, tillering and panicle initiation growth stages."}
kable(N_rates)

@@

```{r figure-4, fig.cap="Tiller sheath blight severity progress was measured in the 2016 experiment at four assessment points. Sheath blight severity was rated on an unevenly spaced ordinal scale and converted to the mid-point percentage value to calculate the area under the disease progress stairs (AUDPS). Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD)."}
ggplot(subset(AUDPS, YEAR == 2016),
       aes(x = WMGT, y = TShB_percent_AUDPS)) +
  geom_boxplot() +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```

@@

```{r figure-5, fig.cap="Tiller sheath blight incidence per hill in the 2015 and 2016 experiments. Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD). Sheath blight incidence was calculated as the number of tillers per hill that were infected with sheath blight divided by the whole number of tillers in the hill. Differing inoculation techniques were used in the 2015 and 2016 experiments, leading to a higher rate of incidence in 2016."}
ggplot(RAW_data, aes(x = WMGT, y = TShB_incidence)) +
  geom_boxplot(fill = NA) +
  theme(axis.text.x = element_text(size = 8,
                                   angle = 45,
                                   hjust = 1)) +
  xlab("Treatment") +
  ylab("Incidence per Hill") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black")) +
  facet_grid(. ~ YEAR)
```

>>>>

The plot design was a split-plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split-plot.

The main plot treatments were alternate wetting and drying (AWD) and continuously flooded (FLD) or farmers' practice as the control treatment. The main plot size was 12m x 12m (144 sq m), with a split-plot size of 5m x 5m (25 sq m). Replication size was 12m x 24m (288 sq m) with a buffer of 1m per split-plot for a whole experiment size of 1,152 sq m.

@@

The split-plot treatment rates of nitrogen were N0 (no nitrogen supply), N100 (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N120 (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) (Table \@ref(tab:N-rates)).

@@

Split-plot N rates differed from the 2015 rates with only two nitrogen treatments, N60 (60 kg per ha as urea split into two applications) and N180 (180 kg per ha in three splits) being applied  (Table \@ref(tab:N-rates)).

@@

In both experiments we were unable to detect any significant effect of AWD on ShB that would cause increases in the disease, which could hinder adoption of the technology. In only one instance were we able to detect any effect of AWD on ShB (Figure \@ref(fig:TShB-sev)). In the 2016 experiment the tiller ShB severity was significantly lower for AWD than the traditional FLD treatment, which may indicate a possible adverse effect of using AWD on tiller ShB severity.

The changes made for the 2016 experiment appear to have improved the experiment, increasing the incidence and decreased variability of sheath blight in the plots (Figure \@ref(fig:TShB-inc)).

@@

```{r N-rates}

kable(N_rates,
      caption = "Nitrogen application rates for 2015 and 2016. Five different nitrogen treatments were applied depending on the experiment. In the 2015 experiment three nitrogen rate treatments were applied: no N (N0), 100 kg per ha N (N100) and 180 kg per ha N (N120). In the 2016 experiment, two nitrogen rate treatments were applied: 60 kg per ha N (N60) and 180 kg per ha N (N180). Treatments were applied in splits, basal, tillering and panicle initiation growth stages.")

@@

```{r TShB-sev, fig.cap="Tiller sheath blight severity progress was measured in the 2016 experiment at four assessment points. Sheath blight severity was rated on an unevenly spaced ordinal scale and converted to the mid-point percentage value to calculate the area under the disease progress stairs (AUDPS). Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD)."}
ggplot(subset(AUDPS, YEAR == 2016),
       aes(x = WMGT, y = TShB_percent_AUDPS)) +
  geom_boxplot() +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```

@@

```{r TShB-inc, fig.cap="Tiller sheath blight incidence per hill in the 2015 and 2016 experiments. Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD). Sheath blight incidence was calculated as the number of tillers per hill that were infected with sheath blight divided by the whole number of tillers in the hill. Differing inoculation techniques were used in the 2015 and 2016 experiments, leading to a higher rate of incidence in 2016."}
ggplot(RAW_data, aes(x = WMGT, y = TShB_incidence)) +
  geom_boxplot(fill = NA) +
  theme(axis.text.x = element_text(size = 8,
                                   angle = 45,
                                   hjust = 1)) +
  xlab("Treatment") +
  ylab("Incidence per Hill") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black")) +
  facet_grid(. ~ YEAR)
```
